---
title: "Model Building"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
#packages
packages <- c("car", "tidyverse", "MASS", "AICcmodavg", "MuMIn", "corrgram", "GGally", "bootStepAIC", "broom","xtable")
source(".\\R_Scripts\\Install_Load_Packages.R")
load_packages(packages)

```

```{r setup, include=FALSE}
# read in data
wolfkde <- read.csv(here::here("Data","wolfkde5.csv"), header = TRUE)
wolfused <-read.csv(here::here("Data","wolfused.csv"), header = TRUE)
wolfavail <-read.csv(here::here("Data","wolfavail.csv"), header = TRUE)
wolfkde_lab3 <- rbind(wolfused, wolfavail)
```

Write out the steps in Model Building 

We're trying to build a conceptual model for modeling wolf habitat use and selection to define habitat for wolves in Colorado 


Steps in Model Building:

1. Look at the data and build a conceptual model of how probability of use is affected by different variables (univariate effects)

2. Go through the steps to build a model
- write out variables of interest and how they affect probability of use by wolves 
- screen for colinearity 
- watch for influential points and outliers
- build one for each pack and for the packs combined

3. Lab 6: Evaluate your model's predictive powers 

# Part 1: Univariate Regression and Colinearity Screening

### Univariate logistic regression
```{r Fitting Univariate Models}
# effect of elevation
elev_univmod <- glm(used ~ Elevation2, 
                    family = binomial(logit),
                    data = wolfkde)
summary(elev_univmod)
# to transform coefficients into odds ratio:
exp(coefficients(elev_univmod))
# Line 400 in Lab 3zx0
# plot it
ggplot(wolfkde, aes(x=Elevation2, y=used)) + geom_point() + stat_smooth(method="glm", method.args = list(family="binomial")) + labs(title = "Elevation Univariate Log Reg")


# effect of human access
human_univmod <- glm(used ~ DistFromHumanAccess2, 
                    family = binomial(logit),
                    data = wolfkde)
summary(human_univmod)
# to transform coefficients into odds ratio:
exp(coefficients(human_univmod))
# plot it
ggplot(wolfkde, aes(x=DistFromHumanAccess2, y=used)) + geom_point() + stat_smooth(method="glm", method.args = list(family="binomial")) + labs(title = "Human Access Univariate Log Reg")

# effect of high human access
highhuman_univmod <- glm(used ~ DistFromHighHumanAccess2, 
                    family = binomial(logit),
                    data = wolfkde)
summary(highhuman_univmod)
# to transform coefficients into odds ratio:
exp(coefficients(human_univmod))
# plot it
ggplot(wolfkde, aes(x=DistFromHighHumanAccess2, y=used)) + geom_point() + stat_smooth(method="glm", method.args = list(family="binomial")) + labs(title = "High Human Access Univariate Log Reg")

# ungulate H.S.I. models
# Sheep
sheep <- glm(used ~ sheep_w2, family=binomial(logit), data=wolfkde)
summary(sheep)
# get odds ratio
exp(coef(sheep))
# plot it
ggplot(wolfkde, aes(x=sheep_w2, y=used)) + geom_point() + stat_smooth(method="glm", method.args = list(family="binomial")) + labs(title = "Sheep Univariate Log Reg")

# Deer
deer <- glm(used ~ deer_w2, family=binomial(logit), data=wolfkde)
summary(deer)
# get odds ratio 
exp(coef(deer))
#2.85 times - strong effects, wolf prob of use increases by 2.85 for each increse in HSI for deer
# plot it
ggplot(wolfkde, aes(x=deer_w2, y=used)) + geom_point() + stat_smooth(method="glm", method.args = list(family="binomial")) + labs(title = "Deer Univariate Log Reg")

# Elk
elk <- glm(used ~ elk_w2, family=binomial(logit), data=wolfkde)
summary(elk)
exp(coef(deer))
# plot it
ggplot(wolfkde, aes(x=elk_w2, y=used)) + geom_point() + stat_smooth(method="glm", method.args = list(family="binomial")) + labs(title = "Elk Univariate Log Reg")

# Moose
moose <- glm(used ~ moose_w2, family=binomial(logit), data=wolfkde)
summary(moose)
exp(coef(moose))
# plot it
ggplot(wolfkde, aes(x=moose_w2, y=used)) + geom_point() + stat_smooth(method="glm", method.args = list(family="binomial")) + labs(title = "Moose Univariate Log Reg")

# Goat
goat <- glm(used ~ goat_w2, family=binomial(logit), data=wolfkde)
summary(goat)
exp(coef(goat))
# plot it
ggplot(wolfkde, aes(x=goat_w2, y=used)) + geom_point() + stat_smooth(method="glm", method.args = list(family="binomial")) + labs(title = "Goat Univariate Log Reg")
```

```{r Making a Coefficient Table}
# make a table of coefficients from the model 
models = rbind(summary(elev_univmod)$coefficients[,1:2], summary(human_univmod)$coefficients[,1:2], summary(highhuman_univmod)$coefficients[,1:2], summary(sheep)$coefficients[,1:2], summary(goat)$coefficients[,1:2], summary(elk)$coefficients[,1:2], summary(moose)$coefficients[,1:2], summary(deer)$coefficients[,1:2])
# Name your models
modelnames = c("elev","disthha", "distacc", "sheep", "goat", "elk", "moose", "deer")
# Now put all of your estimates in a pretty table with names that you'll remember!
estimates.all = matrix(models, nrow=2*length(modelnames), ncol=2, dimnames = list(paste(rep(modelnames, each=2),c("intercept", "coefficient")), c("B", "SE")))
# Gives you a table with b and standard error
xtable(estimates.all)
# export it as a latec - Jordan is presenting on flextable in the R group this semester 
# xtable and flextable allows you to export tables to word
plot(estimates.all)

```

```{r Fitting Categorical Models}
# logistic regression with categorical variables:
wolfkde[,"landcov.f"] <- as.factor(wolfkde$landcov.f)
class(wolfkde$landcov.f)
#469
contrasts(wolfkde$landcov.f) = contr.treatment(13)

# Do logistic regression with categories
oc = glm(used~I(landcov.f=="Open Conifer"),data=wolfkde, family = binomial(logit))
summary(oc)

# Now do the full model 
full = glm(used~I(landcov.f), data=wolfkde, family = binomial(logit))
summary(full)
```


### Test for colinearity and confounding
- generally a rho from cor.test of 0.3 to 0.7 is too confounded
-look for unstable coefficients: a coefficient is unstable when the sign changes or there is a change in 20% (or in ecology more like 50%) of the coefficient from the logistic regression when in the presence of multiple variables

Do you keep in just the variables that are stable?


```{r}
# testing for colinearity 

cor.test(wolfkde$Elevation2, 
         wolfkde$DistFromHighHumanAccess2)
# correlation of .529 means that they are about twice as likely to be the same 
# these variables are also confounded 

# Look for colinearity
summary(elev_univmod)$coefficients[,1:2]
# coefficient is -0.005
summary(highhuman_univmod)$coefficients[,1:2]
# coefficient is -0.0002
elev_highhuman <- glm(used ~ Elevation2 + DistFromHighHumanAccess2, family = binomial(logit),data = wolfkde)
summary(elev_highhuman)$coefficients[,1:2]
# coefficient for elevation is -0.007
# coefficient for distance to high human access is 0.0002: sign flipping
# coefficient for elevation is more stable than the distance to high human access 
# what about distance to human access????????????

# Screening between elk and deer
# Just deer model
summary(deer)$coefficients[,1:2]
# coefficient: 1.11
# Just elk model
summary(elk)$coefficients[,1:2]
# coefficient: 1.12
deer_elk <- glm(used~deer_w2 + elk_w2, data =wolfkde, family= binomial(logit))
# combined model 
summary(deer_elk)$coefficients[,1:2]
# deer coef: .74
# elk coef: .63
cor.test(wolfkde$deer_w2, wolfkde$elk_w2)


# test categorical variables
cor.test(wolfkde$alpine, wolfkde$Elevation2)
cor.test(wolfkde$burn, wolfkde$Elevation2)
cor.test(wolfkde$closedConif, wolfkde$Elevation2)
cor.test(wolfkde$herb, wolfkde$Elevation2)
cor.test(wolfkde$goat_w2, wolfkde$Elevation2)

# checking multiple at the same time
cor.prob2 <- function(X, dfr = nrow(X) - 2) {
  R <- cor(X, use="complete.obs")
  above <- row(R) < col(R)
  r2 <- R[above]^2
  Fstat <- r2 * dfr / (1 - r2)
  R[above] <- 1 - pf(Fstat, 1, dfr)
  Rstar = ifelse(R[above]<0.05, "***", "NS")
  R[above]=paste(R[above],Rstar)
  R
}
# test between dummy variables and other variables
test <- cor.prob2(as.matrix(wolfkde[,c("Elevation2", "DistFromHumanAccess2", "openConif", "closedConif", "modConif", "burn", "herb", "decid", "burn", "alpine")]))
# test between biotic and main variables
test2 <- cor.prob2(as.matrix(wolfkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2")]))
# test between all of them
test3 <- cor.prob2(as.matrix(wolfkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2","openConif", "closedConif", "modConif", "burn", "herb", "decid", "burn", "alpine")]))


cor.prob(as.matrix(wolfkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2")]))
# which outputs to be worried about? Anything with an absolute value greater than .3 to .5
```

There are no locations that are far from from human access and also at low elevation - group these together

### testing all correlations at once 
ggcorr: If they're too correlated (rho > .3), they can't be in the same model 

vif: A value greater than 10 is an indication of potential multi-collinearity problems. As a rough guideline, values between 1 and 2 are not a major problem, but values approaching 10 should cause concern.


```{r}
# Test colinearity of numeric data 
# ggcorr - I like this one best
# can only use this with numeric data - use it to pick out which values to investigate whether they should be removed
ggcorrplot1 <- ggcorr(wolfkde[1:9], label = TRUE)
# check output for values that are over .3 



full.model = glm(used~deer_w2 + elk_w2 +moose_w2 +sheep_w2+goat_w2+Elevation2+DistFromHumanAccess2+DistFromHighHumanAccess2, data =wolfkde, family= binomial(logit))
# no error message about colinearity 
summary(full.model)
# check if your coefficients are stable or bouncing 

vif(full.model)
#line 395 in Lab 5

```



```{r Old Code}
# corrgram 
# corrgram(wolfkde[1:9], order=TRUE, lower.panel=panel.shade,
#          upper.panel=panel.pie, text.panel=panel.txt,
#          main="Correlations in the Wolf Data")

# ggpair
# ggpairplot<-ggpairs(wolfkde[1:9])
# ggpairplot


# 
# cor.prob2 <- function(X, dfr = nrow(X) - 2) {
#   R <- cor(X, use="complete.obs")
#   above <- row(R) < col(R)
#   r2 <- R[above]^2
#   Fstat <- r2 * dfr / (1 - r2)
#   R[above] <- 1 - pf(Fstat, 1, dfr)
#   Rstar = ifelse(R[above]<0.05, "***", "NS")
#   R[above]=paste(R[above],Rstar)
#   R
# }
# 
# cor.prob2(as.matrix(wolfkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2")]))

```


# Using Interactions

You can use interactions to break up or account for colinearity 

Colinear variables of concern: 
-elevation and distance to high human access
-