---
title: "Model_Building_Combined_Rescale_Rasters"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---

```{r Read Packages, include = FALSE}
#packages
packages <- c("car",
              "tidyverse", 
              "MASS", 
              "AICcmodavg", 
              "MuMIn", 
              "corrgram", 
              "GGally", 
              "bootStepAIC", 
              "broom",
              "xtable",
              "sf",
              "terra",
              "tmap")
source(".\\R_Scripts\\Install_Load_Packages.R")
load_packages(packages)

```

# 0.5 Read in Data
```{r Read Data, include=FALSE}
# read in data
wolfkde <- read.csv(here::here("Data","wolfkde5.csv"), header = TRUE)
```

This data has had burn and alpine combined. Justification is that wolves are likely to treat all burned habitat similarly (find a source), and they are also likely to treat all alpine habitat similarly (find a source). 

# Scale Variables

Why do we scale the variables?
*You need to do this so that you can directly compare the coefficients.*

```{r Scaling}
wolfkde2 <- wolfkde
wolfkde2$deer_w2 <- scale(wolfkde$deer_w2)
wolfkde2$moose_w2 <- scale(wolfkde$moose_w2)
wolfkde2$elk_w2 <- scale(wolfkde$elk_w2)
wolfkde2$sheep_w2 <- scale(wolfkde$sheep_w2)
wolfkde2$goat_w2 <- scale(wolfkde$goat_w2)
wolfkde2$Elevation2 <- scale(wolfkde$Elevation2)
wolfkde2$DistFromHumanAccess2 <- scale(wolfkde$DistFromHumanAccess2)
wolfkde2$DistFromHighHumanAccess2 <- scale(wolfkde$DistFromHighHumanAccess2)
#wolfkde <- wolfkde %>% select(-wolf_w2)
# not working?
# now rename them 
#names(wolfkde2) <- gsub("[,1]","", names(wolfkde2))
#wolfkde2 %>% rename(deer_w2=deer_w2[,1],moose_w2=moose_w2[,1])
```

# Separate by Pack

```{r}
wolfkde2 <- na.omit(wolfkde2)
# remove human access
bvkde <- wolfkde2 %>% filter(pack == "Bow Valley")
rdkde <- wolfkde2 %>% filter(pack == "Red Deer")

nrow(bvkde)
# how many used points for each pack
nrow(rdkde %>% filter(used ==0))
```

# Step 1: Screen for Colinear and Unstable Variables and Step 2: Investigate Univariate Relationships

```{r}
# Remove everything with an absolute value greater than 0.5 
## Justification: .3 cuts out too many variables 
source(".\\R_Scripts\\Testing_Colinearity.R")
# look at the colinear variables and look into which one has the lowest AIC value of the univariate comparisons
source(".\\R_Scripts\\One_Univariate_Logistic_Regression.R")


# For bow valley
cortest_bv <- cor.prob2(as.matrix(bvkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2","openConif", "closedConif", "modConif",   "decid", "mixed", "herb", "shrub", "water", "rockIce", "burn", "alpine")]))
# regen gave nothing: leave that out
# The ones that are too correlated:
## #Deer and Elk
## #Elk and Moose 
## Elevation and Deer
## #Elevation and Elk
## #Elevation and Moose
## Elevation and Goat
## Deer with dist high human access
## #Elk with dist high human access
# rockIce with elk, deer, and moose
# Elevation and rockIce
## Elevation and dist to high human access - also confounded

# comparison 1: Human Access vs High Human Access
logreg(bvkde, bvkde$DistFromHumanAccess2, "Distance From Human Access")
# AIC 1161
logreg(bvkde, bvkde$DistFromHighHumanAccess2,"Distance From High Human Access")
# AIC 1115 
# Keep: High Human Access

# comparison 2: elk, deer, and moose
logreg(bvkde, bvkde$deer_w2, "Deer HSI")
# AIC 1108
logreg(bvkde, bvkde$elk_w2, "Elk HSI")
# AIC 1134
logreg(bvkde, bvkde$moose_w2, "Moose HSI")
# AIC 1413, weakest effect
# Keep: Deer

# comparison 3: elevation and distance to high human access 
## Justification - can't keep both of these in bc confounding as well as colinear
logreg(bvkde, bvkde$Elevation2, "Elevation")
# AIC 863
logreg(bvkde, bvkde$DistFromHighHumanAccess2, "High Human Access")
# AIC 1115 

# comparison 4: rockIce with deer/moose/elk (just compare it with deer since we decided to leave that one out)
logreg(bvkde, bvkde$rockIce, "Rock and Ice")
# AIC 1372
logreg(bvkde, bvkde$deer_w2, "Deer")
# 1108
# Keep Deer

# Starting models: 
# 1. With Deer, High Human Access, Goat, Sheep, Land Cover EXCEPT rockIce
#2. With Deer, Sheep, Landcover EXCEPT rockIce


## Cant have elevation with deer? start over with this model with elevation removed
# Elevation model explains it better - try kicking out deer from this full model 


# For red deer
cortest_rd <- cor.prob2(as.matrix(rdkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2","openConif", "closedConif", "modConif",   "mixed", "herb", "shrub", "water", "rockIce", "burn", "alpine")]))
# decid and regen have no values, remove
# Elk, deer and moose
# Elk and sheep
# Elevation and deer/elk moose
# Elevation and dist from human access (not high)
# Distance from human access and distance from high human access not colinear but we should still not include them in the model since they're essentially the same measurement
# Deer/Elk/Moose and rock Ice
# Elevation and rock ice

# Comparison 1: deer/elk/moose
logreg(rdkde, rdkde$deer_w2, "Deer HSI")
# AIC 476
logreg(rdkde, rdkde$elk_w2, "Elk HSI")
# AIC 465
logreg(rdkde, rdkde$moose_w2, "Moose HSI")
# AIC 470
# Keep: Elk

# Comparison 2: Elk and sheep
logreg(rdkde, rdkde$sheep_w2, "Sheep HSI")
# AIC 525, keep Elk

# Comparison 3: high human access and human access
logreg(rdkde, rdkde$DistFromHumanAccess2, "Distance From Human Access")
# AIC 459
logreg(rdkde, rdkde$DistFromHighHumanAccess2,"Distance From High Human Access")
# AIC 529 
# Keep distance to human access - why is this one so much worse?

# Comparison 4: elk and elevation
# Elk AIC 465
logreg(rdkde, rdkde$Elevation2, "Elevation")
# AIC 425
# elevation has a lower AIC but we're going to keep Elk into the model that excludes distance to human access and kicks out elevation

# Comparison 5: elk and Rock ice
# Elk AIC 465, coefficient 1
logreg(rdkde, rdkde$rockIce, "Rock and Ice")
# AIC 440, coefficient -4
# can't kick out all of the prey base, keep elk in the model and kick out rock ice 

# Comparison 6: elevation and rock ice 
# Rock ice 440
# AIC Elevation 425

# Starting models
#1: elk, sheep, goat, distance to human access, land cover EXCEPT rockIce
#2: sheep, goat, elevation, land cover EXCEPT rockIce




# choose elk based on the fact that it makes up most of their diet??? Look into Hebblewhite paper on this
```



## Building/fitting the model
- Use AIC or BIC to determine the best model with stepwise model selection
- use dredging for the models once you've removed colinear variables

use dredging with BIC 

LINE 609
```{r}
# Top models Bow Valley
bv1Top_dredge <- glm(used~deer_w2+
                       DistFromHighHumanAccess2 + 
                       closedConif + 
                       modConif + 
                       sheep_w2 + 
                       goat_w2, 
                     data = bvkde, 
                     family=binomial(logit))

bv3Top_dredge<- glm(used~Elevation2+
                      sheep_w2+
                      modConif+
                      burn, 
                    data=bvkde, 
                    family=binomial(logit))

bv3Top_dredge2 <- glm(used~Elevation2+
                        sheep_w2+
                        burn, 
                      data=bvkde, 
                      family=binomial(logit))
# justification: selected the top scoring AIC model from the step selection models and the top scoring dredbe models by BIC scores. One of the top scoring models from dredge was also the top forward model
# dropped the deer only model and kept elevation because the BIC was ~150 points lower with elevation
# the ones dropped were poor at describing the variation (AIC and BIC was higher than the second model)
```

```{r}
# Top Red Deer models
rd1Top_dredge1 <- glm(used~elk_w2+
                        DistFromHumanAccess2+
                        burn+
                        goat_w2+
                        openConif,
                      data=rdkde,
                      family=binomial(logit))

rd2Top_dredge <- glm(used ~ Elevation2+ 
                       burn+  
                       sheep_w2+ 
                       mixed+ 
                       openConif, 
                     data=rdkde, 
                     family=binomial(logit))

rd2Top_forward <-glm(used ~ Elevation2+ 
                       burn+ 
                       sheep_w2+ 
                       openConif+ 
                       mixed+ 
                       shrub+ 
                       goat_w2+ 
                       herb, 
                     family = binomial(logit), 
                     data = rdkde)

# justification: didn't include any stepAIC models from model 1 because the VIFs were problematic
# justification: the two dredge models had close BICs so I need to validate them to determine which is best

# Elevation was colinear with deer/elk/moose so wasn't included in this model 
```


# Validating the model
```{r Model Validation}
prey_lancov <- rast(".\\Data\\tif_files\\lab6Stack.tif")
bvkde$usedFactor <-as.factor(bvkde$usedFactor)
rdkde$usedFactor <-as.factor(rdkde$usedFactor)

# cross validation
source(".\\R_Scripts\\kxv.R")
# this code splits your data into 5 partitions and validates the data witholding each piece

# Kfolds with a 'fuzz' factor
kxvPrintFlag=FALSE
kxvPlotFlag=TRUE
# this is for breaking ties - add a 1% difference in the case of a tie
kxvFuzzFactor = 0.01
```


```{r K Folds Bow Valley}
# Bow Valley Models
kf_bv1Dredge <- kxvglm(bv1Top_dredge$formula, data=bvkde, k=5, nbin=10) 
kf_bv1Dredge
# deer and distance to high human access
# BIC 931.3

kf_bv3Dredge <- kxvglm(bv3Top_dredge$formula, data=bvkde, k=5, nbin=10) 
kf_bv3Dredge
# elevation
# BIC 857.1

# smoother increase and BIC
# want high rho values - this means more of the variation in our data is explained by the model
# want low p value - more significant 

kf_bv3Dredge2 <- kxvglm(bv3Top_dredge2$formula, data=bvkde, k=5, nbin=10) 
kf_bv3Dredge2
# BIC 857.3
# THIS IS THE TOP MODEL FOR BOW VALLEY
# justification: they had comparable BIC values, but the top dredge 2 one had a more monotomic increase in the k folds cross validation

bvTop <- glm(used~Elevation2+
                        sheep_w2+
                        burn, 
                      data=bvkde, 
                      family=binomial(logit))

summary(bvTop)
confint(bvTop, method= "boot")


# take a look at the confidence intervals 
confint(bvTop, method= "boot")
# justification: we use this because standard errors in a linear regression model don't translate to a confidence interval

```

```{r K Folds Red Deer}
# Start here - look at the differences between and choose a top model 
kf_rd1Dredge1 <- kxvglm(rd1Top_dredge1$formula, data=bvkde, k=5, nbin=10) 
kf_rd1Dredge1
# BIC is 439
# This is the top model
# justification: better monotomic increase with habitat quality

kf_rd2Dredge <- kxvglm(rd2Top_dredge$formula, data=bvkde, k=5, nbin=10) 
kf_rd2Dredge
# BIC: 416.1

kf_rd2Forward <- kxvglm(rd2Top_forward$formula, data=bvkde, k=5, nbin=10) 
kf_rd2Forward
# AIC 397.2
hist(wolfkde$deer_w2)

rdTop <- glm(used~elk_w2+
                        DistFromHumanAccess2+
                        burn+
                        goat_w2+
                        openConif,
                      data=rdkde,
                      family=binomial(logit))
summary(rdTop)

# take a look at the confidence intervals 
confint(rdTop, method= "boot")


```

With these outputs, I am looking for an increasing monotomic relationship between habitat quality and frequency of use

# Spatial prediction
```{r Quick Plots}
par(mfrow = c(1,1)) # reset graphical parameters
# fit the model
bvkde$fitted_top <- fitted(bvTop)
# Graph the output
ggplot(bvkde, aes(EASTING, NORTHING, col = fitted_top)) + geom_point(size=5) + coord_equal() +  scale_colour_gradient(low = 'yellow', high = 'red')

rdkde$fitted_top <- fitted(rdTop)
ggplot(rdkde, aes(EASTING, NORTHING, col = fitted_top)) + geom_point(size=5) + coord_equal() +  scale_colour_gradient(low = 'yellow', high = 'red')
```

```{r Raster Wrangling}
# Plot the spatial predictions
kernelHR <- st_read(".\\Data\\shapefiles\\homerangeALL.shp")

wolfyht <- st_read(".\\Data\\shapefiles\\wolfyht.shp")

# maske a mask raster
kernels <- rast()
ext(kernels) <- c(xmin=546836, xmax=612093, ymin=5662036, ymax=5748911) 

# load in full raster from Lab6
full_rast <- rast(".\\Data\\tif_files\\lab6Stack.tif")
full_rast <- scale(full_rast, center=TRUE, scale=TRUE)
#full_rast <- dropLayer(full_rast,c(6,10:19))

# names(full_rast)
# plot(full_rast$DEER_W) # good
# plot(full_rast$MOOSE_W, main = "Moose") # good
# plot(full_rast$ELK_W, main = "Elk") # good
# plot(full_rast$SHEEP_W, main = "Sheep") # good
# plot(full_rast$GOAT_W, main = "Goat") # good
# plot(full_rast$WOLF_W, main = "Wolf") # good
# plot(full_rast$Elevation2, main = "Elevation") # good I think
# plot(full_rast$DistFromHumanAccess2, main = "Dist From Human Access") # good I think
# plot(full_rast$DistFromHighHumanAccess2, main = "Dist From High Human Access") # good I think
# plot(full_rast$HABITATTYPE, main = "Habitat Type") # error with this one

# read in landcover data
# only use the landcover layers that you need
landcover <- rast(".\\Data\\tif_files\\landcover16.tif")
#plot(landcover)
cats(landcover)
#par(mar=c(1,1,1,1))

# separate out each land cover type and make the extent match the full raster
alpine <- ifel(landcover == 15 | landcover == 16, 1,ifel(is.na(landcover),NA,0))
#alpine <- crop(alpine,full_rast,snap="near")

burn <- ifel(landcover == 12 | landcover == 13 |landcover == 14, 1, ifel(is.na(landcover),NA,0))
ext(burn)
#burn <- resample(burn,full_rast,method="near")
full_rast <- project(full_rast,burn)
openConif <- ifel(landcover == 1, 1, ifel(is.na(landcover),NA,0))
openConif <- resample(openConif,full_rast,method="near")

closedConif <- ifel(landcover == 3, 1, ifel(is.na(landcover),NA,0))
herb <- ifel(landcover == 7, 1, ifel(is.na(landcover),NA,0))
mixed <- ifel(landcover == 5, 1, ifel(is.na(landcover),NA,0))
rockIce <- ifel(landcover == 10, 1, ifel(is.na(landcover),NA,0))
water <- ifel(landcover == 9, 1, ifel(is.na(landcover),NA,0))
modConif <- ifel(landcover == 2, 1, ifel(is.na(landcover),NA,0))
decid <- ifel(landcover == 10, 1, ifel(is.na(landcover),NA,0))
```


```{r Spatial Prediction Bow Valley}
## Pull out the coefficients
bvTop_coefs <- bvTop$coefficients

elevation <- full_rast$Elevation2
sheep <- full_rast$SHEEP_W
# need to center and scale each raster

# put through each raster
plot(full_rast$Elevation2)
plot(bvkde, add = TRUE)


rast_bvTop <- (exp(bvTop_coefs[1]+
                    bvTop_coefs[2]*full_rast$Elevation2+
                    bvTop_coefs[3]*full_rast$SHEEP_W+
                     bvTop_coefs[4]*burn))/(1+exp(bvTop_coefs[1]+
                    bvTop_coefs[2]*full_rast$Elevation2+
                    bvTop_coefs[3]*full_rast$SHEEP_W+
                    bvTop_coefs[4]*burn))

# exp(bvTop_coefs[1]+
#                     bvTop_coefs[2]*5000+
#                     bvTop_coefs[3]*6+
#                      bvTop_coefs[4]*1)/(1+exp(bvTop_coefs[1]+
#                     bvTop_coefs[2]*5000+
#                     bvTop_coefs[3]*6+
#                     bvTop_coefs[4]*1))
# 
# rast_bvTop <- (exp(bvTop_coefs[1]+
#                     bvTop_coefs[2]*elevation+
#                     bvTop_coefs[3]*sheep+
#                      bvTop_coefs[4]*burn))/(1+exp(bvTop_coefs[1]+
#                     bvTop_coefs[2]*elevation+
#                     bvTop_coefs[3]*sheep+
#                     bvTop_coefs[4]*burn))
#                     
cats(rast_bvTop)
plot(rast_bvTop)
# plot your predictions
plot(rast_bvTop, col=colorRampPalette(c("yellow","orange","red"))(255), ext=kernels)
plot(kernelHR, add=TRUE, col = NA)
plot(wolfyht, col='blue', pch = 16, add=TRUE)
#Line 814



```

```{r Spatial Prediction Red Deer}
## Pull out the coefficients
rdTop_coefs <- rdTop$coefficients

# put through each raster

rast_rdTop <- exp(rdTop_coefs[1]+
                    rdTop_coefs[2]*full_rast$ELK_W+
                    rdTop_coefs[3]*full_rast$DistFromHumanAccess2+
                    rdTop_coefs[4]*burn+
                    rdTop_coefs[5]*full_rast$GOAT_W+
                    rdTop_coefs[6]*openConif)/(1+exp(rdTop_coefs[1]+
                    rdTop_coefs[2]*full_rast$ELK_W+
                    rdTop_coefs[3]*full_rast$DistFromHumanAccess2+
                    rdTop_coefs[4]*burn+
                    rdTop_coefs[5]*full_rast$GOAT_W+
                    rdTop_coefs[6]*openConif))
cats(rast_bvTop)
plot(rast_rdTop, col=colorRampPalette(c("yellow","orange","red"))(255), ext=kernels)
plot(kernelHR, add=TRUE, col = NA)
plot(wolfyht, col='blue', pch = 16, add=TRUE)
# follows use points well 
#Line 814


```


```{r Model Averaging}
# how to do model averaging?
model_avg <- terra::app(c(rast_rdTop,rast_bvTop),mean)
plot(model_avg, col=colorRampPalette(c("yellow","orange","red"))(255), ext=kernels)
plot(kernelHR, add=TRUE, col = NA)
plot(wolfyht, col='blue', pch = 16, add=TRUE)

# looks like it's predicting our data well except for where it is in the corner - does this have high human access?
plot(full_rast$DistFromHumanAccess2)
# this is becuase bow valley model is mostly driven by elevation

```


he wants the biological reasoning behind things 
dropping elevation? 
- this is colinear with human to access 

- do one for each pack and then do model averaging with spatial prediction

at the end, we'll need to create a raster stack 

Validate your spatial predictions with 10 equal area bins and leave one out then put wolf telemetry data back over it and see how well it works 
