---
title: "Model_Building_Combined"
author: "Anna Kurtin"
date: "`r Sys.Date()`"
output: html_document
---

```{r Read Packages, include = FALSE}
#packages
packages <- c("car",
              "tidyverse", 
              "MASS", 
              "AICcmodavg", 
              "MuMIn", 
              "corrgram", 
              "GGally", 
              "bootStepAIC", 
              "broom",
              "xtable",
              "sf",
              "terra",
              "tmap")
source(".\\R_Scripts\\Install_Load_Packages.R")
load_packages(packages)

```

# 0.5 Read in Data
```{r Read Data, include=FALSE}
# read in data
wolfkde <- read.csv(here::here("Data","wolfkde5.csv"), header = TRUE)
```

This data has had burn and alpine combined. Justification is that wolves are likely to treat all burned habitat similarly (find a source), and they are also likely to treat all alpine habitat similarly (find a source). 

# Scale Variables

Why do we scale the variables?
*You need to do this so that you can directly compare the coefficients.*

```{r Scaling}
wolfkde2 <- wolfkde
wolfkde2$deer_w2 <- scale(wolfkde$deer_w2)
wolfkde2$moose_w2 <- scale(wolfkde$moose_w2)
wolfkde2$elk_w2 <- scale(wolfkde$elk_w2)
wolfkde2$sheep_w2 <- scale(wolfkde$sheep_w2)
wolfkde2$goat_w2 <- scale(wolfkde$goat_w2)
wolfkde2$Elevation2 <- scale(wolfkde$Elevation2)
wolfkde2$DistFromHumanAccess2 <- scale(wolfkde$DistFromHumanAccess2)
wolfkde2$DistFromHighHumanAccess2 <- scale(wolfkde$DistFromHighHumanAccess2)
#wolfkde <- wolfkde %>% select(-wolf_w2)
# not working?
# now rename them 
#names(wolfkde2) <- gsub("[,1]","", names(wolfkde2))
#wolfkde2 %>% rename(deer_w2=deer_w2[,1],moose_w2=moose_w2[,1])
```

# Separate by Pack

```{r}
wolfkde2 <- na.omit(wolfkde2)
# remove human access
bvkde <- wolfkde2 %>% filter(pack == "Bow Valley")
rdkde <- wolfkde2 %>% filter(pack == "Red Deer")
```

# Step 1: Screen for Colinear and Unstable Variables and Step 2: Investigate Univariate Relationships

```{r}
# Remove everything with an absolute value greater than 0.5 
## Justification: .3 cuts out too many variables 
source(".\\R_Scripts\\Testing_Colinearity.R")
# look at the colinear variables and look into which one has the lowest AIC value of the univariate comparisons
source(".\\R_Scripts\\One_Univariate_Logistic_Regression.R")


# For bow valley
cortest_bv <- cor.prob2(as.matrix(bvkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2","openConif", "closedConif", "modConif",   "decid", "mixed", "herb", "shrub", "water", "rockIce", "burn", "alpine")]))
# regen gave nothing: leave that out
# The ones that are too correlated:
## #Deer and Elk
## #Elk and Moose 
## Elevation and Deer
## #Elevation and Elk
## #Elevation and Moose
## Elevation and Goat
## Deer with dist high human access
## #Elk with dist high human access
# rockIce with elk, deer, and moose
# Elevation and rockIce
## Elevation and dist to high human access - also confounded

# comparison 1: Human Access vs High Human Access
logreg(bvkde, bvkde$DistFromHumanAccess2, "Distance From Human Access")
# AIC 1161
logreg(bvkde, bvkde$DistFromHighHumanAccess2,"Distance From High Human Access")
# AIC 1115 
# Keep: High Human Access

# comparison 2: elk, deer, and moose
logreg(bvkde, bvkde$deer_w2, "Deer HSI")
# AIC 1108
logreg(bvkde, bvkde$elk_w2, "Elk HSI")
# AIC 1134
logreg(bvkde, bvkde$moose_w2, "Moose HSI")
# AIC 1413, weakest effect
# Keep: Deer

# comparison 3: elevation and distance to high human access 
## Justification - can't keep both of these in bc confounding as well as colinear
logreg(bvkde, bvkde$Elevation2, "Elevation")
# AIC 863
logreg(bvkde, bvkde$DistFromHighHumanAccess2, "High Human Access")
# AIC 1115 

# comparison 4: rockIce with deer/moose/elk (just compare it with deer since we decided to leave that one out)
logreg(bvkde, bvkde$rockIce, "Rock and Ice")
# AIC 1372
logreg(bvkde, bvkde$deer_w2, "Deer")
# 1108
# Keep Deer

# Starting models: 
# 1. With Deer, High Human Access, Goat, Sheep, Land Cover EXCEPT rockIce
#2. With Deer, Elevation, Sheep, Landcover EXCEPT rockIce


# For red deer
cortest_rd <- cor.prob2(as.matrix(rdkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2","openConif", "closedConif", "modConif",   "mixed", "herb", "shrub", "water", "rockIce", "burn", "alpine")]))
# decid and regen have no values, remove
# Elk, deer and moose
# Elk and sheep
# Elevation and deer/elk moose
# Elevation and dist from human access (not high)
# Distance from human access and distance from high human access not colinear but we should still not include them in the model since they're essentially the same measurement
# Deer/Elk/Moose and rock Ice
# Elevation and rock ice

# Comparison 1: deer/elk/moose
logreg(rdkde, rdkde$deer_w2, "Deer HSI")
# AIC 476
logreg(rdkde, rdkde$elk_w2, "Elk HSI")
# AIC 465
logreg(rdkde, rdkde$moose_w2, "Moose HSI")
# AIC 470
# Keep: Elk

# Comparison 2: Elk and sheep
logreg(rdkde, rdkde$sheep_w2, "Sheep HSI")
# AIC 525, keep Elk

# Comparison 3: high human access and human access
logreg(rdkde, rdkde$DistFromHumanAccess2, "Distance From Human Access")
# AIC 459
logreg(rdkde, rdkde$DistFromHighHumanAccess2,"Distance From High Human Access")
# AIC 529 
# Keep distance to human access - why is this one so much worse?

# Comparison 4: elk and elevation
# Elk AIC 465
logreg(rdkde, rdkde$Elevation2, "Elevation")
# AIC 425
# elevation has a lower AIC but we're going to keep Elk into the model that excludes distance to human access and kicks out elevation

# Comparison 5: elk and Rock ice
# Elk AIC 465, coefficient 1
logreg(rdkde, rdkde$rockIce, "Rock and Ice")
# AIC 440, coefficient -4
# can't kick out all of the prey base, keep elk in the model and kick out rock ice 

# Comparison 6: elevation and rock ice 
# Rock ice 440
# AIC Elevation 425

# Starting models
#1: elk, sheep, goat, distance to human access, land cover EXCEPT rockIce
#2: sheep, goat, elevation, land cover EXCEPT rockIce




# choose elk based on the fact that it makes up most of their diet??? Look into Hebblewhite paper on this
```



## Building/fitting the model
- Use AIC or BIC to determine the best model with stepwise model selection
- use dredging for the models once you've removed colinear variables

use dredging with BIC 

LINE 609
```{r}
# Top models Bow Valley
bv2Top_backward <- glm(formula = used ~ deer_w2 +
                         Elevation2 + 
                         sheep_w2 + 
                         modConif + 
                         burn, 
                       family = binomial(logit),
                       data = bvkde)
bv2Top_dredge <- glm(used~deer_w2+
                       Elevation2+
                       burn+
                       modConif+
                       sheep_w2, 
                     data=bvkde,
                     family=binomial(logit))

# justification: took out the first dredge model because the BIC was ~100 higher
# justification: only used second model because the first was poor at describing the variation (AIC and BIC was higher than the second model)
```

```{r}
# Top Red Deer models
rd1Top_dredge1 <- glm(used~elk_w2+
                        DistFromHumanAccess2+
                        burn+
                        goat_w2+
                        openConif,
                      data=rdkde,
                      family=binomial(logit))

rd2Top_dredge <- glm(used ~ Elevation2+ 
                       burn+  
                       sheep_w2+ 
                       mixed+ 
                       openConif, 
                     data=rdkde, 
                     family=binomial(logit))

rd2Top_forward <-glm(used ~ Elevation2+ 
                       burn+ 
                       sheep_w2+ 
                       openConif+ 
                       mixed+ 
                       shrub+ 
                       goat_w2+ 
                       herb, 
                     family = binomial(logit), 
                     data = rdkde)

# justification: didn't include any stepAIC models from model 1 because the VIFs were problematic
# justification: the two dredge models had close BICs so I need to validate them to determine which is best
```


# Validating the model
```{r Model Validation}
prey_lancov <- rast(".\\Data\\tif_files\\lab6Stack.tif")
bvkde$usedFactor <-as.factor(bvkde$usedFactor)
rdkde$usedFactor <-as.factor(rdkde$usedFactor)

# cross validation
source(".\\R_Scripts\\kxv.R")
# this code splits your data into 5 partitions and validates the data witholding each piece

# Kfolds with a 'fuzz' factor
kxvPrintFlag=FALSE
kxvPlotFlag=TRUE
# this is for breaking ties - add a 1% difference in the case of a tie
kxvFuzzFactor = 0.01

# Bow Valley Models
kf_bv1_backwards = kxvglm(bv2Top_backward$formula, data=bvkde, k=5, nbin=10) 
kf_bv1_backwards
## Start here 3/8
## Lab 6 line 700
```

ROC

he wants the biological reasoning behind things 
dropping elevation? 
- this is colinear with human to access 

- do one for each pack and then do model averaging with spatial prediction

at the end, we'll need to create a raster stack 

Validate your spatial predictions with 10 equal area bins and leave one out then put wolf telemetry data back over it and see how well it works 



```{r Code Graveyard}
# for combined data
# source(".\\R_Scripts\\Run_Univariate_Logistic_Regressions.R")
# # Strength of univariate variables for bow valley wolves
# univ_logregs(bvkde)
# # Strength of univariate variables for red deer wolves 
# #univ_logregs(rdkde)
# # Strength of univariate variables for both packs 
# #univ_logregs(wolfkde)
# 
# 
# #check out interesting relationships/visualize it
# source(".\\R_Scripts\\Plot_Univariate_Regressions.R")
# make_plot(data = wolfkde, 
#           xaxis = wolfkde$goat_w2,
#           title = "Goat Univariate Regression")
# 
# 
# # For combined wolf data 
# cortest1 <- cor.prob2(as.matrix(wolfkde[,c("deer_w2","elk_w2", "moose_w2", "sheep_w2", "goat_w2", "Elevation2", "DistFromHumanAccess2", "DistFromHighHumanAccess2","openConif", "closedConif", "modConif", "burn", "herb", "decid", "burn", "alpine")]))

# The values that we are concerned with in this model are
## Elk, deer and moose are all collinear 
## Elevation with elk, deer, moose, and goat
## Elevation and disance to human access and high human access are confounded
## distance to human access and deer, elk, and moose


### Considering interactions
 
# Only include interactions between the effects of a categorical and a continuous variable
# 
# # first convert the categorical variable into a factor
# wolfkde$closedConif <- as.factor(wolfkde$closedConif)
# ggplot(wolfkde, aes(x=DistFromHumanAccess2, y=used, fill=closedConif)) + stat_smooth(method="glm", method.args = list(family="binomial"), level=0.95) #+ facet_grid(closed~.)
# 
# Note, there is a bit stronger evidence of an interaction here (the lines cross), which brings us back to our original observation above. 
```

